Using News' Title and Body Text Features to Detect Fake News: a Viability Study

### 2019-09-05 FakeNewsAnalysis/paperLatex/paperFelipeNevesDataScience.tex
finished paper

The goal of this study was to define the viability of using features extracted from both titles
and body from news article for the detection of Fake News using a large dataset and comparing
the results with these available on previous bibliography.
\end{abstract}

@@

Automatic Fake News detection is a topic of growing interest in the last couple of years,
especially due to the influences presented in the 2016 American Elections \cite{allcott_gentzkow_2017},
however it is still very difficult to detect this type of content and stop its spread,
specially in social networks and messaging apps unregulated environments. These situations
alone give enough reason to pursue the definition of reliable approaches to Fake News detection.

@@

Besides the features already mentioned in \ref{table:horne_features} an additional set of features
was used for the evaluating the performance of the algorithm, these features are listed on Table
\ref{table:expanded_features} and will be referred as 'expanded feature set' whereas the features from
Table \ref{table:horne_features} will be referred as 'original feature set'. In the same way the dataset
from Horne's work \cite{horne_2017} will be referred as 'original dataset' and the dataset from Szpakowski
\cite{szpakowski_2019} wil be referred as 'expanded dataset'.

@@

As it is clear to see in Figure \ref{fig:results} it is clear that the performance over
this bigger dataset was significantly worse than in the original paper \cite{horne_2017},
however it is important to understand why whe had such difference.

@@

However only using more features is good but may have some drawbacks in regarding the processing
speed that will be increased. Also, if the features set contains features that are not relevant
to the analysis the performance may actually decrease due to containing inputs that cannot be
separated by the classifier, such as what happened with the performance of out classification using
the original and expanded features over the titles of the original dataset, where the performance
dropped from 72.67\% with the original features to 71.33\% when using the expanded features, as shown
on Table \ref{table:od_of} and Table \ref{table:od_ef} respectively.

>>>>

In the past few years there have been a large number of algorithms developed to detect
Fake News. We analyze the viability of using features extracted from both titles and
body from news article for the detection of Fake News using a large dataset and comparing
the results with these available on previous bibliography. We find that features used
to classify a one dataset may not be appropriate to classify from a different one, due
to the differences presented in these datasets. At last we test our model achieving a
classification accuracy of 73.93\%.

\end{abstract}

@@

Automatic Fake News detection is a topic of growing interest in the last couple of years,
especially due to the its impact in the 2016 American Elections \cite{allcott_gentzkow_2017}\cite{dey_rafi_parash_arko_chakrabarty_2018},
however it is still very difficult to detect this type of content and stop its spread,
specially in social networks and messaging apps unregulated environments. These situations
alone give enough reason to pursue the definition of reliable approaches to Fake News detection.

@@

Besides the features already mentioned in Table \ref{table:horne_features} an additional set of features
was used for the evaluating the performance of the algorithm, these features are listed on Table
\ref{table:expanded_features} and will be referred as 'expanded feature set' whereas the features from
Table \ref{table:horne_features} will be referred as 'original feature set'. In the same way the dataset
from Horne's work \cite{horne_2017} will be referred as 'original dataset' and the dataset from Szpakowski
\cite{szpakowski_2019} wil be referred as 'expanded dataset'.

@@

As it is clear to see in Figure \ref{fig:results} it is clear that the performance over
this bigger dataset was significantly worse than in the original paper \cite{horne_2017},
however it is important to understand why we had such difference.

@@

However only using more features is good but may have some drawbacks in regarding the processing
speed that will be increased. Also, if the features set contains features that are not relevant
to the analysis the performance may actually decrease due to containing inputs that cannot be
separated by the classifier, such as what happened with the performance of out classification using
the original and expanded features over the titles of the original dataset, where the performance
dropped from 72.67\% with the original features to 71.33\% when using the expanded features, as shown
on Table \ref{table:od_of} and Table \ref{table:od_ef} respectively. Simply adding more features
will not necessarily cause an improvement in the performance \cite{volkova_shaffer_jang_hodas_2017}.
